---
title: 'My Amazing Solutions to Problem Set 5: Intro to Longitudinal Data'
author: "ANONYMOUS"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r setup, message=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(arm)
library(knitr)
library(texreg)
library(stargazer)
library(lmtest)
library(dplyr)
library(ggeffects)
library(foreign)
opts_chunk$set(message=FALSE, warning=FALSE, tidy=FALSE, out.width="4.5in")
options(digits=3, width = 60)
# clear workspace
rm(list = ls())

#read in dataset 
dat <- read.dta("~/Desktop/Fall 2023/Multilevel-HW/PS 5/alcuse.dta")
```


# The tasks
a. Substitute the level—2 models into the level—1 model to obtain the reduced form model.

Reduced model:
$$ Y_{tj} = \gamma_{00} + \gamma_{01} w_{1j} + \gamma_{02} w_{2j} + u_{0j} + \gamma_{10}time_{tj} + \gamma_{11} w_{1j}time_{tj} + \gamma_{12}time_{tj}+ w_{2j}time_{tj} + u_{1j} + time_{tj} + \epsilon_{tj} $$

The level 1 model
$$ Y_{tj} = \pi_{0j} + \pi_{1j} time_{tj} + \epsilon_{tj} $$

and the level 2 models
$$
\begin{aligned}
\pi_{0j} &= \gamma_{00} + \gamma_{01} w_{1j} + \gamma_{02} w_{2j} + u_{0j} \\
\pi_{1j} &= \gamma_{10} + \gamma_{11} w_{1j} + \gamma_{12} w_{2j} + u_{1j}
\end{aligned}
$$
with $(u_{0j}, u_{1j})$ bivariate normal as usual.]

\bigskip

b. Determine the correct `lmer()` call and fit the model.

``` {r linear growth curve}
mod.b <- lmer(alcuse ~ 1 + coa*age_14 + peer*age_14 + (1 + age_14|id), data = dat)

display(mod.b)
```

\bigskip

c. *Succinctly* identify and interpret all the parameter estimates. No more than one sentence per parameter. Make a list, with each item being, e.g., "$\gamma_{00}$ = #, which means blah, blah blah."  

1.$\gamma_{00}$ = -0.32, which means at age 14, the frequency of alcohol use for a typical child is -0.32. 

2.$\gamma_{01}$ = 0.58, we would expect adolescents with a parent diagnosed with alcohol abuse to have 0.58 unites higher frequency of alcohol use as compared to adolescents without an alcohol dependent parent, holding all else equal. 

3.$\gamma_{02}$ = 0.69,  we would expect adolescents with one unit increase in alcohol use among peers to have 0.69 unites higher frequency of alcohol use.

4.$\gamma_{10}$ = 0.43, we would expect a type adolescent to have increased 0.43 units of frequency of alcohol use per year. 

5.$\gamma_{11}$ = -0.01, the growth rate of alcohol use frequency is expected to be 0.01 units lower among adolescents with a parent diagnosed with alcohol abuse compared to adolescents without a parent diagnosed with alcohol abuse.

6.$\gamma_{12}$ = -0.15, the growth rate of alcohol use frequency is expected to be 0.15 units lower among adolescents with one unit higher in alcohol use among peers.

7.$\sqrt\tau_{00}$ = 0.51, standard deviation of the random intercept

8.$\sqrt\tau_{11}$ = 0.39, standard deviation of the random slope 

9.$\hat\sigma$ = 0.58, residual variation within adolescents. 

\bigskip

d. Is there evidence that children of parents diagnosed with alcohol dependency are different than the other children?  Formally test and then explain any differences.  

We will perform a t-test on the coefficient term for the interaction between age 14 and coa. The null hypothesis is that $\gamma_{11}$ = 0 

```{r}
summary(mod.b, correlation = F)
# get the confidence intervals 
fixef(mod.b) - 1.96 * se.fixef(mod.b)
fixef(mod.b) + 1.96 * se.fixef(mod.b)

```

We see that the growth rate of alcohol use frequency is expected to be 0.014 (95%CI: -0.263, 0.2351, p-value = 0.912 > 0.05) units lower among adolescents with a parent diagnosed with alcohol abuse compared to adolescents without a parent diagnosed with alcohol abuse.We would fail to reject the null (95% CI captures the null value of 0 and pvalue is greater than 0.05), saying that there isn't enough evidence that children of parents diagnosed with alcohol dependency are different than the other children. 

\bigskip

e. Make an 80% confidence interval for the effect of peer use on the growth rate of a kid. Using average growth as a benchmark, interpret both the top and bottom end of your confidence interval in terms of being a large or small effect.

```{r confidence interval}
fixef(mod.b) - 1.28 * se.fixef(mod.b)
fixef(mod.b) + 1.28 * se.fixef(mod.b)
``` 

We are 80% confident that the average adolescent with one unit higher in peer alcohol use has between 0.472 and 0.9170 units higher in alcohol use frequency. 

At the bottom end of the CI, the difference in alcohol use frequency for adolescents with one unit higher in peer alcohol use is 0.472, which is medium. 

At the upper end of the CI, the difference in alcohol use frequency for adolescents with one unit higher in peer alcohol use is 0.9170, which is large. 

\bigskip

f. Plot your data.  You can either plot the individual growth curves of all your children, use ``ggeffects`` to make a plot of growth for children of parents diagnosed with alcohol dependency vs. the other children, or some other plot of your choosing that shows your model results.^[See the online textbook for sample code to make these plots.]

```{r plot, fig.width=6, fig.height=3}
dat <- dat |> 
  mutate(linear_fit = fitted(mod.b))

# plot 
ggplot(dat, aes(x = age_14, y = linear_fit, color = factor(coa), group = factor(id))) +
  geom_line() +
  geom_point(aes(y = alcuse)) +
  labs(x = "Age in years since first interview",
       y = "Alcohol use frequency", 
       color = "children of parents diagnosed with alcohol dependency") +
  theme(legend.position = "bottom")

ggeffect(mod.b, terms = c("age_14", "coa")) |>
plot()



# Create the ggplot
ggplot(dat, aes(x = age_14, y = linear_fit)) +
  # Plot the line for linear_fit
  geom_line(aes(group = 1), color = "blue") +  # Added group and color for clarity
  # Plot the points for alcuse
  geom_point(aes(y = alcuse, color = "Alcohol Use")) +  # Added color legend for points
  # Add labels and title
  labs(
    x = "Age in years since first interview",
    y = "Alcohol use frequency",
    title = "Individual Curves for Alcohol Use Frequency"  # Added a title
  ) +
  # Customize theme
  theme(
    legend.position = "bottom",
    legend.title = element_text("Legend"),  # Added legend title
    plot.title = element_text(hjust = 0.5)  # Centered the plot title
  )


```
\bigskip

g. Using the empirical Bayes estimates predictions from your model, estimate the fraction of the study participants who decreased their use of alcohol over the course of the study.

```{r }
display(mod.b)

0.43 - 3*0.12 # still positive
```
We know that 99.7% of the data fall within 3SD of mean. And 3sd below the average is still slightly positive (0.07). Therefore, we might say that less than 0.3% of the study participants has decreased use of alcohol over the course of the study. 

\bigskip

h. We can estimate the _reliability_ of our empirical Bayes estimates, which is a measure of how well we think we are measuring the estimate. The reliability is calculated as

$$ reliability = 1 - \frac{\mbox{uncertainty in estimate}}{\mbox{uncertainty in est} + \mbox{variance of effects}} = 1 - \frac{SE^2_j}{SE^2_j + \tau }$$
where $\tau$ is the variance of the random effects being evaluated (e.g., it would be $\tau_{00}$ if we wanted the reliability of the intercepts and $\tau_{11}$ if we wanted the reliability of the slopes).

What is the reliability of the estimated growth rates in your model?  Does this make you trust or not trust your estimates in part (d), above?^[Usually a reliability of 0.80 or so is considered "reliable" and less than that we have to account for the fact that our estimates are in large part noise.]  

```{r reliability}
reliability = 1- (0.12^2/(0.12^2 + 0.39))
reliability 
```

The reliability is 0.964 and this makes me trust my estimated growth rates. 
