---
title: 'PS 1: My amazing solutions'
output:
  pdf_document: default
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center" )

# Secret code (e.g., loading packages and data) goes here.

library( tidyverse )
library(readstata13) # This lets us read STATA files!
library(lme4)
#you can add any additional libraries here

```

## 1. **Tenn STAR, an experiment**

### a) Looking at the standard errors 

The standard errors did/did not change after accounting for clustering....


### b) Extend your model
```{r}
# some code goes here!
```

Small class size did/did not have an impact on test scores....


### c) Further control for classroom characteristics

```{r}
# some code goes here!
```

Explain what happened when you included fixed effects for classroom...


### d) Testing a hypothesis about teachers' years of experience and small class sizes

```{r}
# some code goes here!
```


\newpage
## 2. **Aggregation practice (coding)**

### a) Make a summarized dataframe with the average attainment, the proportion male, and the number of observations for each neighborhood in the dataset.
```{r}
# load the data and summarize it
dat <- read.dta13("neighborhood.dta")

# summarize the data
dat_sum <- dat %>% 
  group_by(neighid) %>%
  summarise(
    mean_attain = mean(attain), 
    prop_male = mean(male), 
    obs = n()) 

# add a variable for deprivation
dat_sum$deprive <- dat$deprive[match(dat_sum$neighid, dat$neighid)]

head(dat_sum)
```

### b) Make a histogram of the proportion male. Is this what you would expect? Is anything odd with this histogram?
```{r}
# make a histogram of the prop_male variable
ggplot(dat_sum, aes(x = prop_male)) + 
  geom_histogram(binwidth = 0.05) + 
  labs(x = "proportion of males", y = "count of neighborhoods")
```

This histogram is odd because there are many neighborhoods that have 0 males and some neighborhoods only have males. This is not what we would expect. 

### c) calculate the average, standard deviation, and range of the number of observations in a neighborhood. What does this tell you about (b)?
```{r}
# calculate the average, standard deviation, and range of the number of observations in a neighborhood
mean(dat_sum$obs)
sd(dat_sum$obs)
range(dat_sum$obs)

# install.packages("gtsummary")
library(gtsummary)

# Present the results in a table
dat_sum %>%
  select(obs) %>%
  tbl_summary(
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{min}, {max}", "{mean} ({sd})"),
    missing = "no", 
    digits = all_continuous() ~ 4
  )
```

We see that the average number of observations is 4.4, the standard deviation is 2.8, and the range is 1 to 16. This tells us that there are some neighborhoods that only have 1 observation. This is why we see the odd histogram in (b).

### d) Scatterplot of the social-deprivation score (deprive) and average attainment. What can you learn, if anything, from this plot? What things are distorted, and how?

```{r}
# make a scatterplot of the social-deprivation score (deprive) and average attainment
ggplot(dat_sum, aes(x = deprive, y = mean_attain)) + 
  geom_point() +
  labs(x = "Social-deprivation score", y = "Average attainment")

  mean(dat_sum$deprive) # average social-deprivation score is 0.37
  range(dat_sum$deprive) # range of social-deprivation score is -1.082  2.959
  mean(dat_sum$mean_attain) # average attainment of the neighborhood averages is 0.08
  range(dat_sum$mean_attain) # average attainment of the neighborhood averages is -1.3276, 2.4151
```
Neighborhoods with higher social deprivation score tend to have lower average attainment. However, there are a few neighborhoods with high social-deprivation scores (above 1) that have high average attainment.
Neighborhoods with the lowest average attainment of -1.3276 have a wide range of social-deprivation scores, which is odd.
The plot is distorted partially because there are some neighborhoods that only have 1 observation, so the average attainment is the same as the attainment of that observation.


