---
title: 'PS 1: My amazing solutions'
output:
  pdf_document: default
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center" )

# Secret code (e.g., loading packages and data) goes here.

library( tidyverse )
library(readstata13) # This lets us read STATA files!
library(lme4)
#you can add any additional libraries here

```

## 1. **Tenn STAR, an experiment**

### a) Looking at the standard errors 

The standard errors did/did not change after accounting for clustering....


### b) Extend your model
```{r}
# some code goes here!
```

Small class size did/did not have an impact on test scores....


### c) Further control for classroom characteristics

```{r}
# some code goes here!
```

Explain what happened when you included fixed effects for classroom...


### d) Testing a hypothesis about teachers' years of experience and small class sizes

```{r}
# some code goes here!
```


\newpage
## 2. **Aggregation practice (coding)**

### a) Summarize the data frame 
```{r}
# load the data 
#dat <- read.dta13("neighborhood.dta")

# some code goes here!
```

### b) Histogram of proportion male
```{r}
# some code goes here!
```

This histogram is (not?) odd because ... 

### c) Observations per neighborhood
```{r}
# some code goes here!
```

We see that ... 

### d) Scatterplot of attainment and social deprivation 
```{r}
# some code goes here!
```

comment on scatter plot...


