---
title: 'Problem Set 6: Longitudinal wtsa and Three-Level Models'
author: "S-043 (Fall 2023)"
wtse: 'Template'
header-includes:
output: pdf_document
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# libraries
library(tidyverse)
library(lme4)
library(lmerTest)
library(arm)
library(knitr)
library(texreg)
library(stargazer)
library(haven)

opts_chunk$set(message=FALSE, warning=FALSE, echo = TRUE, out.width="4.5in")
options(dplyr.summarise.inform = FALSE, digits = 3)

# set ggplot theme
theme_set(theme_classic())

# clear memory
rm(list = ls())
## You can use this block if you want to do your wtsa cleaning in your Markdown file. You are also welcome to do your wtsa cleaning in an entirely separate script. 
# Loading the wtsa
wts <- read_dta("/Users/Betty/Desktop/Fall 2023/Multilevel-HW/PS 6 /childweights.dta")

glimpse(wts)
summary(wts)
```

# Linear and Quadratic Growth Models

## 1a. Plot
```{r results = 'hide'}
set.seed(123)

# Load the ggplot2 library
library(ggplot2)

# Select 16 random children based on "id"
selected_ids <- sample(wts$id, 16)

# Filter the wtsaset to include only the selected children
selected_wts <- wts[wts$id %in% selected_ids, ]

# Create a scatter plot with growth lines using ggplot2
ggplot(data = selected_wts, aes(x = age, y = weight, group = id)) +
  geom_point() +
  geom_line() +
  labs(x = "Age (years)", y = "Weight (kg)", title = "Weight vs. Age for 16 Randomly Selected Children") +
  theme_minimal()
```

No need to turn in anything here.

## 1b. Fit a Linear Growth Model to these wtsa

```{r}
# centering age 
wts$age = wts$age - mean(wts$age)

LG <- lmer(weight ~ 1 + age + gender + (1 + age|id),data = wts)
display(LG)
```

## 1c. Fit a Quadratic Growth Model

```{r}
wts$agesq = wts$age^2

#The following code will not run
#QG <- lmer(weight ~ 1 + age + agesq * gender + (1 + age + agesq|id), data = wts)

table( wts$id ) # tally up obs/child
table( table( wts$id ) ) # tally the tallys
# Add number of obs for each group to your wtsa.
wts = wts %>% group_by( id ) %>%
mutate( n = n() )
```

Distribution of the number of observations per child: Most children have 2 to 4 observations. A few has 1 or 5 observations, 

1  2  3  4  5 
4 19 27 15  3 
 
## 1d. Subset the data to fit the Quadratic Model

```{r}
# Remove children with only one observation
id_counts <- table( wts$id )

# Find the ids that appear more than once
ids_to_keep <- as.numeric(names(id_counts[id_counts > 1]))

# Filter the dataframe to keep only rows with ids that appear more than once
wts <- wts[wts$id %in% ids_to_keep, ]


QG <- lmer(weight ~ 1 + age + agesq * gender + (1 + age + agesq|id),
data = wts)
display(QG)
```

## 1e. Linear vs. Quadratic

```{r}
display(LG) #6
display(QG) #14

#calculate R-squared 
install.packages("MuMIn")
library(MuMIn)

rsquared <- r.squaredGLMM(QG)

# Print the R-squared value
cat("R-squared:", rsquared, "\n")

rsquared <- r.squaredGLMM(LG)

# Print the R-squared value
cat("R-squared:", rsquared, "\n")

```
The datasets are not the same. Therefore, we cannot use AIC. The models are not nested. Therefore,we cannot use LRT. 
The R-squared for the LG model is slighly lower than the QG model (0.976 vs. 0.857). However, a more complex model will always have higher R-squared. In my opnion, the complex model isn't worth it because we are adding 8 parameters with only about 10% more variation explained by the covariates. In addition, we are lossing data from children who only had one observation

## 1f. Plot the Growth Curves
```{r}
set.seed(123)
# Select 16 random children based on "id"
selected_ids <- sample(wts$id, 16)

# Filter the wtsaset to include only the selected children
selected_wts <- wts[wts$id %in% selected_ids, ]

# Create a scatterplot with growth curves
ggplot(data = selected_wts, aes(x = age, y = weight, group = id, color = factor(gender))) +
  facet_wrap(~ id) +
  geom_point() +
  geom_line(aes(y = predict(QG, newdata = selected_wts), color = factor(gender)), linetype = "dashed") +
  geom_line(aes(y = predict(LG, newdata = selected_wts), color = factor(gender))) +
 labs(title = "Child weight growth over time. Dashed lines are for the quadratic growth model and solid lines are for linear growth model",
x = "Age (months)",
y = "Predicted weight",
caption = "Age is centered at 1.08 years") +
theme(axis.text.x = element_text(size = rel(.75), angle = 00))



```

\newpage

# Piecewise Linear Growth (READS wtsa)

## 2a. Mathematical Model

[Equation here]

## 2b. Answer RQ1

```{r cache = TRUE}
studs = readRDS("./READS_student_data.Rds")

#define a new summer variable as the different between summer and school
studs$summer  = studs$summer - studs$school
m1 <- lmer(score ~ 1 + summer +
(1 + summer|id) +
(1 + summer|sch),
data=studs)
display(m1)
confint(m1)
```
The different in growth rate between summer and school year is statistically significant with 95%CI of -1.816, -1.594.

## 2c. Answer RQ2

```{r cache = TRUE}
m2 <- lmer(score ~ 1 + summer*frl +
(1 + summer|id) + #frl does not vary within students
(1 + summer*frl|sch),
data=studs)

display(m2)
```
At school level, we are including 4 random effects: random effect by free lunch, random effect by growth rate, random intercept and random growth rate by free lunch. 

At students level, since free lunch dose not vary within students, we did not add a random effect by free lunch at the student level. 

## 2d. Concept Check 1

Theoretically, it will be possible. We can still force to add a random effect of FRL as the school level. However, since either almost all or no students receive FRL in a specific school, we will see a very small SD for the random effect of FRL at the school level. We will also be adding unnecessary complexity to our model. M2 has 4 fixed effects, 1 level 1 variance term, 2 level 2 variance terms (and 1 cov), 4 level 3 variance terms (6 covs), summing up to a total of 18 parameters!!

M1 has 2 fixed effects, 1 level 1 variance term, 2 level 2 variance terms (and 1 cov), 2 level 3 variance terms (and 1 cov). Giving us a total of just 9 parameters. 

