---
title: 'My Amazing Solutions for Problem Set 3: Inference and Multilevel Modeling'
author: "ANONYMOUS"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

**This is a group assignment. Please do not include your names in your PDF submission (Canvas group membership connects your submission to your names without the teaching team seeing them!).**

\begin{center}
\huge ~ The Problems ~
\end{center}


1. Present a favorite plot or display from your work
====================================================================== 
 
```{r Q1}
# Your code
```

This plot... 

\newpage

2. Does sense of emotional safety vary by gender and grade?
======================================================================
[Your model here!]

```{r Q2}
#Your code here
```
The model indicates there are/are not gender differences...

\newpage

3. Do those who feel discriminated against feel less safe?
======================================================================
[Your model here!]

```{r Q3}
#Your code
```

The model indicates...


\newpage	


4. Contextual effects for discrimination?
======================================================================
``` {r Q4} 
#your code
``` 

a) Is there a contextual effect for discrimination with regard to emotional safety?  

Yes/No...

b) Once you take out the contextual effect, if any, does the relationship at the student level of sense of discrimination and emotional safety change?  

The relationship does/does not change... 


Once you have answered this, reflect on the following:  
c) What, if anything, does it add substantively to do this within-school and contextual effect analysis?  

It adds/doesn't add...

