---
title: 'My Amazing Solutions for Problem Set 3: Inference and Multilevel Modeling'
author: "Jen, Marshae and Xin"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

**This is a group assignment. Please do not include your names in your PDF submission (Canvas group membership connects your submission to your names without the teaching team seeing them!).**

\begin{center}
\huge ~ The Problems ~
\end{center}

```{r setup, include=FALSE}
# clear memory
rm(list = ls())

#load dataset 
dat = readRDS( "~/Desktop/Fall 2023/Multilevel-HW/PS 3/mycleandata.rds" )
#load packages
library(tidyverse)
library(lme4)
library(lmerTest)
#install.packages("lmerTest")
library(ggplot2)
#install.packages("broom.mixed")
library(broom.mixed)
#install.packages("texreg")
library(texreg)
```

1. Present a favorite plot or display from your work
====================================================================== 
 
```{r Q1}

```

This plot... 

\newpage

2. Does sense of emotional safety vary by gender and grade?
======================================================================
Assuming that slopes for gender and grade is the same within schools and between schools, in the population.
[Your model here!]

```{r Q2}
#standardizing esafe and disc
dat$esafe_std <- scale(dat$esafe)
dat$disc_std <- scale(dat$disc)

#mean centering disc
dat %>% group_by(ID) %>% mutate(mean_disc = mean(disc)) -> dat

dat <- dat %>%
  mutate(disc_dev = disc_std - mean_disc)




```
The model indicates there are/are not gender differences...

\newpage

3. Do those who feel discriminated against feel less safe?
======================================================================
**Answer**: The statistical model should allow the relationship between emotional safety and feeling discriminated to vary across school.
$$
\begin{aligned}
esafe_{ij} &= \beta_{0j} + \beta_{1j} (disc - \overline{disc})_{ij} + \varepsilon_{ij} \\
\beta_{0j} &= \gamma_{00} + \gamma_{01} \overline{disc}_j +\zeta_{0j} \\
\beta_{1j} &= \gamma_{10} + \zeta_{1j}
\end{aligned}
$$
where $i$ indexes students and $j$ indexes schools.
```{r Q3}
m3 <- lmer(esafe_std ~ disc_dev + mean_disc + (disc_dev|ID), dat)
m3 %>% tidy() %>% knitr::kable(digits = 3)
arm::display(m3)
#confidence intervals 
confint(m3, method = c("Wald")) %>% knitr::kable(digits = 3)
```
The model indicates that there is a negative relationship between emotional safety and feeling discriminated against.
On average, a one-unit increase in SD of discrinimation score predicts a 0.36 (95%CI:-0.402, -0.320) units decrease the SD of emotional safety. This is statistically significant (95% CI does not include 0).
Among schools, schools with higher mean discrimination scores have lower mean emotional safety scores. 
On average, a one-unit increase in SD of school mean discrinimation score predicts a 0.87 (95%CI:-0.790, 0.417) units decrease the SD of emotional safety. This is statistically significant (95% CI does not include 0).

\newpage	


4. Contextual effects for discrimination?
======================================================================
``` {r Q4} 
#your code
``` 

a) Is there a contextual effect for discrimination with regard to emotional safety?  

Yes/No...

b) Once you take out the contextual effect, if any, does the relationship at the student level of sense of discrimination and emotional safety change?  

The relationship does/does not change... 


Once you have answered this, reflect on the following:  
c) What, if anything, does it add substantively to do this within-school and contextual effect analysis?  

It adds/doesn't add...

