---
title: 'Problem Set 8: Logistic Growth Curves'
author: "S-043 (Fall 2023)"
date: 'Template'
header-includes:
output: pdf_document
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(lme4)
library(lmerTest)
library(arm)
library(ggsci) # for cool color pallets
library(texreg)

options(digits=3)

# clear workspace 
rm(list = ls())
```

```{r, include=FALSE}
# this is the "initial_analysis" script 

# load cleaning dataset 
# you must run the "prepare_data" script
  dat = readRDS("cleaned_data.rds")
  head( dat )
  
  # aggregate data by time and treatment status
  haz <- dat %>% group_by(EN, time.f, mtime) %>%
    summarise(ch_haz = mean(ch_haz),
              ch_stunted = mean(ch_stunted))
```

## 1. Fit a Logistic Regression Model
Random effects linear growth model with ch_stunted as the outcome, allowing an time*EN interaction, include random intercept sand slopes for both child and LHW 

```{r}
m1<-glmer(ch_stunted~mtime*EN +(1+mtime|ch_id) + (1+mtime|lhw_id), data=dat, binomial, control=glmerControl(optimizer="bobyqa" ))

summary(m1)
```

\newpage

## 2. Interpret the Results of the Model (Answer RQ2)

### 2a. Assess whether EN significantly impacts the initial odds of being stunted.  Does this result make sense?

*Make sure log odds is okay and not transformed to probability *
The initial log odds  of ch_stunting for the treatment effect ENTx at baseline is 0.247, but this is not statistically significant at the 0.05 significance level with a p-value (0.275)>0.05. Beyond the fact that treatment was randomized at baseline, indicating that we would not expect there to be any statistically significant differences at baseline (before initiation). Over time we do see a statistically significant difference, when treatment is interacted with time. 

### 2b. Assess whether EN significantly impacts how the odds of a child being stunted changes over time.

The log odds of stunting over time looking at EN:TIME is -0.065; this  is statistically significant, p-value of 0.001< 0.05 significance level, allowing us to reject the null hypothesis that EN does not change the log odds of ch_stunting outcome across levels of time. This makes sense, as we have hypothesized that the treatment once initiated would aid in decreasing the odds of stunting for those who randomized to receive it, as they age. 

### 2c. Interpret the fixed effect coefficients of the logistic linear growth model in terms of odds multipliers.
```{r}
# odds ratio formulation
exp( fixef( m1 ) )
```
The odds of childhood stunting at baseline occurring is 0.0707 times the odds of it not occurring, when mtime=0 and treatment=0. 
Time odds ratio: 1.24, For each one-unit increase in the mtime variable, the odds of the event childhood stunting occurring increase by ~1.23 times, holding treatment constant. 
Treatment odds ratio: 1.28, The odds of childhood stunting for those who have the treatment is 1.28 times the odds of childhood stunting for those who do not have the treatment (control group), holding time constant.
Interaction time and treatment: Comparing the treated group of children, vs the untreated group of children (controls), the additional difference above and beyond the difference in the odds of stunting across levels of time is 0.937  

*JEN PLEASE CHECK MY PHS BRAIN* ALSO THE 1.28 just doesnt make sense to me, treatment is coded as 1 for treated,0 for untreated right? 
\newpage

## 3. Make and interpret a median growth trajecotry plot

### 3a. Enter in and adapt the following code so it works with your script:

```{r, eval = FALSE}
npd = expand.grid( ch_id = -1,
                     lhw_id = -1, 
                     mtime = seq( 0, 24, by = 1 ),
                     EN = unique( dat$EN ) )
npd$ch_stunted = predict( m1, newdata=npd, 
                            allow.new.levels=TRUE, 
                            type="response" )
ggplot( npd, aes( mtime, ch_stunted, col=EN ) ) +
    geom_line() + geom_point()
```

### 3b. Interpret what this plot shows us in terms of individual or population average impacts. If individual, describe the individual we are seeing. 

Demonstrates the effect of the treatment over time for the predicted probability of being stunted at the population level.  The proportion of children who experience stunting over time, is different for those within treatment, vs those who were in the control group. 
Overall the proportion of stunting overtime, lower height for age z scores increased over time for both groups, but the odds for stunting for those in the treatment group was lower than the odds of stunting in control group over time. 

\newpage

## 4. Plot the aggregate of the individual predictions as a model check

```{r, eval = FALSE}
alldat = expand.grid( ch_id = unique( dat$ch_id ),
                      EN = unique( dat$EN ),
                      mtime = unique( dat$mtime ) )

# Add in the child's LHW 
children = dplyr::select( dat, ch_id, lhw_id ) %>% unique()
alldat = merge( alldat, children, by="ch_id" )

# Predict for all these points
alldat$pr_stunt = predict( m1, newdata=alldat, type="response" )

aggdat = alldat %>% group_by( EN, mtime ) %>%
  summarise( pr_stunt = mean( pr_stunt ) )
```

Using the `aggdat` dataframe, make a plot just like you did in the first part of this assignment.
How does this plot compare to your raw data? Is there evidence of model misfit?

```{r}
library(ggplot2)

# Create the plot
ggplot(aggdat, aes(x = mtime, y = pr_stunt, color = EN)) +
  geom_line() + 
  geom_point() +
  labs(x = "mtime", y = "Predicted Probability of Stunting", color = "EN") +
  theme_minimal()
```

The plot compares relatively well, showing increase in predicted probability of stunting over time, varied by treatment level. Treated predicted probability of stunting over time is lower than that of those in the control over time. Curve appears to be a bit different, for both?? 


\newpage

## 5. A sensitivity check

### Make a regression table

```{r}
# code here
```

### 5a. Did the point estimate change?

[Answer here]

### 5b. Did the SE change?

[Answer here]

### 5c. Did anything else change?

[Answer here]

### 5d. Are any covariates connected to the outcome?

[Answer here]

\newpage

## 6. Write a formal wrap-up

[Answer here]
