---
title: 'My Amazing Solutions to Problem Set 7: Complex Models'
author: "ANONYMOUS"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r initial exploration, include = FALSE}
# Initial Exploration (included here for convenience)
  # This is the code from the script "initial_exploration.R" 
  # Get familiar with this code and what it is doing before you proceed 

options( digits=3 )

# load packages 
library(tidyverse)
library(arm)
library(readstata13)
library(ggsci) # for cool color pallets


dat = readRDS( "cleaned_data.rds" )

head( dat )

# aggregate data by time and tx group.
haz <- dat %>% group_by(EN, time.f, mtime) %>% 
  summarise(ch_haz = mean(ch_haz), 
            ch_stunted = mean(ch_stunted),
            .groups = "drop" )
haz 

# Plot two outcomes by time and treatment group  
ggplot(haz, aes(x=time.f, y=ch_haz, group=EN, color=EN)) +
  geom_line(lwd=1.2) + geom_point() +
  labs(x="Time", y="Average", 
       title="Average HAZ by intervention arm and time point",
       color="Intervention arm") +
  theme(legend.position="bottom") + scale_colour_startrek() 


ggplot(haz, aes(x=time.f, y=ch_stunted, group=EN, color=EN)) +
  geom_line(lwd=1.2) + geom_point() +
  labs(x="Time", y="Proportion", 
       title="Stunting by intervention arm and time point",
       color="Intervention arm") +
  theme(legend.position="bottom") + scale_colour_startrek() 



# Look at population differences in proportion stunted at different time points
# between tx groups at time points
head( haz )
haz.w = haz %>%
  pivot_wider( values_from = c(ch_haz, ch_stunted), 
               names_from = "EN" ) 

head( haz.w )
haz.w = mutate( haz.w, Delta_stunt = ch_stunted_Tx - ch_stunted_Co )
haz.w



## ----Simple multilevel model ------------------------------------------------------------


M1 = lmer(ch_haz ~ mtime * EN  + (1 + mtime|ch_id) + (1|lhw_id), 
              data=dat,
              control = lmerControl(optimizer = "bobyqa"))
display( M1 )


# Look at some of the random effects
qplot( coef(M3)$ch_id[[1]] )
qplot( coef(M3)$lhw_id[[1]] )

```

\newpage 
\begin{center}
\huge ~ The Problems ~
\end{center}

# 1. Some code review questions

a) For the cleaned and prepared data, what is the `EN` variable and how was it made?

b) What was done with children with missing outcome data?

c) What are the `mtime` and `time.f` variables and how are they different?

\newpage 

# 2. Evaluate potential model fit

[Your explanation here]

\newpage 

# 3. Identify a mathematical model

[Your model here]

\newpage 

# 4. Make a prediction interval

```{r prediction interval}

```

[The 95% prediction interval is... There is a large/small amount of individual variation...] 

\newpage 

\begin{center}
\large ~ Extension:\\ Improving Model Fit ~
\end{center}

# 5. Allowing for a curve

There are # parameters... 

\newpage 

# 6. Plot the treatment curves

```{r plots}
# Some plotting code here 


# For convenience, here are some of the code snippets from the assignment sheet:
groups = expand.grid( ch_id = -1,
                     lhw_id = -1, 
                     time.f = unique( dat$time.f ),
                     EN = unique( dat$EN ) )
times = parse_number( levels( dat$time.f ) )
groups$mtime = times[ groups$time.f ]
```

[Your assessment] 

\newpage 

# 7. Use predicted values to create plots for 16 random children

```{r individual plots}
#Some code for plotting the 16 children here 

``` 

[Your further assessment]

\newpage 

# 8. Interpret the results of the model (Answer RQ 1)

[Your analysis addressing RQ1 including: 
  a) Assessing whether EN significantly impacts baseline HAZ.  Does this result make sense?

  b) Assessing whether EN significantly impacts child growth over time.

  c) Does the estimated size of the impact seem notable? E.g., you might consider how much extra does a child grow over the first 24 months?]

\newpage 
