---
title: 'My Amazing Solutions to Problem Set 7: Complex Models'
author: "ANONYMOUS"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r initial exploration, include = FALSE}
# Initial Exploration (included here for convenience)
  # This is the code from the script "initial_exploration.R" 
  # Get familiar with this code and what it is doing before you proceed 

options( digits=3 )

# load packages 
library(tidyverse)
library(arm)
library(readstata13)
library(ggsci) # for cool color pallets


dat = readRDS( "./PS 7/cleaned_data.rds" )

head( dat )

# aggregate data by time and tx group.
haz <- dat %>% group_by(EN, time.f, mtime) %>% 
  summarise(ch_haz = mean(ch_haz), 
            ch_stunted = mean(ch_stunted),
            .groups = "drop" )
haz 

# Plot two outcomes by time and treatment group  
ggplot(haz, aes(x=time.f, y=ch_haz, group=EN, color=EN)) +
  geom_line(lwd=1.2) + geom_point() +
  labs(x="Time", y="Average", 
       title="Average HAZ by intervention arm and time point",
       color="Intervention arm") +
  theme(legend.position="bottom") + scale_colour_startrek() 


ggplot(haz, aes(x=time.f, y=ch_stunted, group=EN, color=EN)) +
  geom_line(lwd=1.2) + geom_point() +
  labs(x="Time", y="Proportion", 
       title="Stunting by intervention arm and time point",
       color="Intervention arm") +
  theme(legend.position="bottom") + scale_colour_startrek() 



# Look at population differences in proportion stunted at different time points
# between tx groups at time points
head( haz )
haz.w = haz %>%
  pivot_wider( values_from = c(ch_haz, ch_stunted), 
               names_from = "EN" ) 

head( haz.w )
haz.w = mutate( haz.w, Delta_stunt = ch_stunted_Tx - ch_stunted_Co )
haz.w

## ----Simple multilevel model ------------------------------------------------------------


M1 = lmer(ch_haz ~ mtime * EN  + (1 + mtime|ch_id) + (1|lhw_id), 
              data=dat,
              control = lmerControl(optimizer = "bobyqa"))
display( M1 )


# Look at some of the random effects
qplot( coef(M3)$ch_id[[1]] )
qplot( coef(M3)$lhw_id[[1]] )

```

\newpage 
\begin{center}
\huge ~ The Problems ~
\end{center}

# 1. Some code review questions

a) For the cleaned and prepared data, what is the `EN` variable and how was it made?
EN is a binary variable, in which "Co" means that enhanced nutrition (EN) was NOT received (i.e, either control or RS) and "Tx" means that EN was received (i.e, either EN or RS + EN). 

```{r}
#table( dat$EN, dat$treatment.f )
```

b) What was done with children with missing outcome data?
Children with missing outcome (missing ch_haz) were dropped.Code that used to drop children with missing outcome is referenced below:  

```{r}
# dat <- filter( dat,
#                !is.na( ch_stunted ),
#                !is.na( ch_haz ) )
```

c) What are the `mtime` and `time.f` variables and how are they different?
Mtime and time.f represent **months since baseline**. For example, mtime=0 means that data was collected at baseline and mtime=6 means that data collected at 6 months since baseline. mtime is a numeric variable, while time.f is a factor variable. 

```{r}
#class(dat$mtime)
#class(dat$time.f)
```

\newpage 

# 2. Evaluate potential model fit

Looking at the average HAZ and proportion of children stunted across survey waves, linear growth model does not seem to be completely appropriate. The trajectory of both average HAZ and stunting seem to be non-linear, where both the average HAZ and proportion of children start to become more steady from 12 month to 24 month. However, fitting linear models may not lead to huge issue because the trajectory are still very roughly linear. 

\newpage 

# 3. Identify a mathematical model

```{r}
M1 = lmer(ch_haz ~ mtime * EN  + (1 + mtime|ch_id) + (1|lhw_id), data=dat,
              control = lmerControl(optimizer = "bobyqa"))
```

$$
\begin{aligned}
y_{ijk} &= \beta_{0jk} + \beta_{1jk} mtime_{tjk} + \varepsilon_{tjk} \\
\\
\varepsilon_{tjk} &\sim N(0,\sigma^2) \\
\beta_{0jk} &= \gamma_{00k} + \mu_{jk} \\
\beta_{1jk} &= \gamma_{10k} + \gamma_{11k} EN_{jk} + \mu_{jk}  \\
\gamma_{00k} &= \alpha_{000} + \nu_k
\\
\begin{bmatrix}
{u_{0i}}\\ 
{u_{1i}}
\end{bmatrix} &~\sim N\left( {\left[ {\begin{array}{*{30}{c}}
{{0}}\\
{{0}} 
\end{array}} \right],\left[ {\begin{array}{*{30}{c}}
{\tau_{00}}&{}\\
{\tau_{10}}&{\tau_{11}}
\end{array}} \right]} \right)\\
\nu_{tjk} &\sim N(0,\sigma_1^2) \\
\end{aligned}
$$
-   Level 1: time, measured in months since baseline
-   Level 2: individual children
-   Level 3: Lady health worker
-   $a_{tjk}$: months since baseline at measurement point *t* for children *j* with lady health worker *k*

\newpage 

# 4. Make a prediction interval
The 95% prediction interval is -3.54 and 1.14. There is a large amount of individual variation.

```{r prediction interval}
# Extract fixed effects
fixed_effects <- fixef(M1)

# Average outcome for children in the control group for median LWH at baseline (mtime = 0 and EN = 0 )
display(M1)
sd_intercept <- 1.17
#random_effects <- ranef(M1)$ch_id
#sd_intercept <- sd(random_effects$`(Intercept)`)
# Calculate predicted outcomes
# For a child with a random intercept 2 SD above the average
predicted_high <- fixed_effects[1] + (2 * sd_intercept)
predicted_high
# For a child with a random intercept 2 SD below the average
predicted_low <- fixed_effects[1] - (2 * sd_intercept)
predicted_low
# The predicted outcomes are for the baseline (intercept) with adjustments for the random effects 2 SD above and below the average.

```
\newpage 

\begin{center}
\large ~ Extension:\\ Improving Model Fit ~
\end{center}

# 5. Allowing for a curve

There are # parameters... 

\newpage 

# 6. Plot the treatment curves

```{r plots}
# Some plotting code here 


# For convenience, here are some of the code snippets from the assignment sheet:
groups = expand.grid( ch_id = -1,
                     lhw_id = -1, 
                     time.f = unique( dat$time.f ),
                     EN = unique( dat$EN ) )
times = parse_number( levels( dat$time.f ) )
groups$mtime = times[ groups$time.f ]
```

[Your assessment] 

\newpage 

# 7. Use predicted values to create plots for 16 random children

```{r individual plots}
#Some code for plotting the 16 children here 

``` 

[Your further assessment]

\newpage 

# 8. Interpret the results of the model (Answer RQ 1)

[Your analysis addressing RQ1 including: 
  a) Assessing whether EN significantly impacts baseline HAZ.  Does this result make sense?

  b) Assessing whether EN significantly impacts child growth over time.

  c) Does the estimated size of the impact seem notable? E.g., you might consider how much extra does a child grow over the first 24 months?]

\newpage 
